Automated Program Verification and Synthesis of Embedded \& Cyber-physical Systems: Advances and Challenges

### 2017-08-07 paper.tex
additions to section 2

% As a general rule, do not put math, special symbols or citations
% in the abstract
\begin{abstract}
Dependency on the correct operation of embedded systems is rapidly growing, mainly due to their wide range of applications, such as micro-grids, automotive device control, health care, surveillance, mobile devices, telecommunications, internet of things, and consumer electronics. Their structures are becoming more and more complex and now require multi-core processors with scalable shared memory, signal-processing pipelines, and sophisticated software modules, in order to meet increasing computational power, flexibility demands, and adaptation to new scenarios and behaviors. Additionally, interaction with real-world entities and addition of modern communication capabilities further enhance the mentioned features and give rise to a new class of systems: embedded \& cyber-physical systems (ECPS). As a consequence, reliability of ECPS becomes a key issue during system development, which must be carefully addressed and assured. Normally, state-of-the-art verification methodologies for ECPS generate test vectors (with constraints) and use assertion-based verification and high-level processor models, during simulation. Nonetheless, other additional challenges have been raised: the need for meeting time and energy constraints, handling concurrent software, dealing with platform restrictions, evaluating implementation-structure choices, validating operation logic, ensuring correct system behavior together with physical plants, providing compliance with target systems, incorporating knowledge about new problems and conditions, and supporting new software architectures and legacy designs. The present research discusses challenges, problems, and recent advances to ensure correctness and timeliness regarding ECPS. Reliability issues, in the development of ECPS, are then considered, as a prominent verification and synthesis application for achieving a \textit{correct-by-construction} design.
\end{abstract}

@@

The interaction between embedded software and physical processes created a different class of systems, which are complex, highly integrated, and present a mixture of continuous and discrete dynamics, named as \textit{hybrid systems} (HSs). Indeed, the embedded-system domain allied to the most recent communication revolution caused an intense integration between different and spread physical processes, which are called embedded \& cyber-physical systems (ECPSs). In particular, the presence of modern communication technologies ({\it e.g.}, internet of things - IoT) causes a revolution in terms of flexibility, scalability, and complexity of those systems and adds a novel class of problems to the challenges found in the embedded systems domain. For example, micro-grids, {\it i.e.}, small-scale electricity systems that gather different sources of distributed generation (DG) and loads, are emerging ECPSs, where reliability and carbon emission reduction are of paramount importance~\cite{xu15}. One may also notice that each DG element and the majority of current loads already employ embedded software and present several safety-critical requirements; however, the integration of both constitute a ECPS, where additional challenges related to synchronization, stability, control, communication, and reliability of the entire micro-grid arise.

@@

Figure~\ref{intelligent-product} shows some examples of ECPS, which typically consist of a human-machine interface ({\it e.g.}, keyboard and display), a processing unit ({\it e.g.}, real-time computer system), and an instrumentation interface ({\it e.g.}, sensor, network, and actuator) that can be connected to some physical plant~\cite{Kopetz11}. Indeed, many current ECPS, such as unmanned aerial vehicles (UAVs)~\cite{groza2015formal} and medical monitoring systems~\cite{Cordeiro09}, become interesting solutions only if they can reliably perform their target tasks. For instance, UAVs are a trend on military missions, due to the absence of pilots; however, an incorrect plan execution may cost civilian lives, which is unacceptable. In addition, wrong disease diagnosis or condition-evaluation reports have the potential to compromise patients' health and even many aspects of their lives, with serious consequences. On the one hand, portable ECPS are capable of monitoring and identifying conditions that are very difficult for a physician, mainly when his contact with patients happens only in medical clinics. On the other hand, wrong or incomplete diagnostic data may delay necessary treatments or deteriorate patients' health. 
%
\begin{figure}[!t]
	\centering
	\includegraphics[scale=0.4]{figure1.jpg}
	\caption{An embedded system is part of a well-specified larger one (intelligent product).}
	\label{intelligent-product}
\end{figure}

Besides, when physical interaction with the real world is needed, which happens in ECPS, additional care must be taken, mainly when human action is directly replaced, as in vehicle driving. Regarding the latter, even human-in-the-loop feedback control can be employed \cite{munir}, which then raises deeper concerns related to reliability of human behavior modeling and system implementation. Consequently, it is important to go beyond design correctness and also address behavior correctness, which may be performed by incorporating system models. Specifically, such models can be used to evaluate and even synthesize a given particular system, by ensuring that all needed functions are correctly implemented and the correct behavior is exhibited, {\it i.e.}, the system is indeed correct by its method of construction~\cite{Abate17}.

@@

The goals of the present article is to briefly present part of the established research and also clarify current trends, which are mainly driven by ECPSs, IoT, and ML. We present a brief discussion about models for ECPS, in Section \ref{sec:model}, and then cover preliminaries on bounded model checking, \textit{k}-induction verification based on invariant inference, Craig interpolants, counterexample guided inductive synthesis, and system model incorporation, in Section~\ref{Preliminaries}. In Section~\ref{Verification-Challenges}, the main challenges regarding verification and synthesis of embedded systems are described, while Section~\ref{Research-Problem} tackles those challenges that are (partially) open in current published research. Section~\ref{achievements} then describes current achievements and future trends in verification and synthesis of ECPS and Section \ref{Newapp} addresses new applications, which take into account the new trends ({\it e.g.}, ML) mentioned here. Finally, we conclude and describe future work, in Section~\ref{conclusions}.

@@

Automated formal verification have been employed to ensure the correctness and reliability of ECPS during the last decade, as well as controller synthesis for ECPS and HS \textcolor{red}{add citations}.

A ECPS $S$ can be represented as a tuple $S=(X,X_{0},U,Y,r,v)$, where $X$ is a set of reachable states, $X_{0}$ is a set of initial states, $U$ is a set of inputs, $Y$ is the set of outputs, $r$ is a transition function such that $r:X\times U \rightarrow X$, and  $v$ is an output mapping function such that $v:X\times U \rightarrow Y$~\cite{Rungger16,Alur00,Girard11}.

The overall automated formal synthesis process for ECPS and HS involves three main steps: \textit{modeling}, which is related to obtain a sound mathematical representation able to capture all relevant aspects of continuous and discrete dynamics and their interaction, as well as, formulate properties and invariants that describe the desirable behavior; \textit{verification}, where the ECPS behavior is checked through the proposed mathematical structure and the attendance of the desired behavior is verified; and, finally, the controller \textit{synthesis} itself, that produces a controller or finds invariant control laws that allow the ECPS to meet the requirements and behave as desired, which is confirmed by a verification oracle.

@@

Models are largely employed in science and engineering areas, in order to analyze, predict, and modify the behavior of real-world systems. Specially, deterministic and stochastic models are the basis of many recent engineering advances. The former is uniquely determined by parameter values and previous states, while the latter presents inherent randomness, which is described by probability distributions~\cite{stochmodel}. Nonetheless, one can argue that such models are not enough to represent ECPSs, since interaction among different entities (software and hardware components) presents several events, which could not be deterministically or stochastically modeled; they can be better represented by nondeterministic approaches~\cite{leeCPS2}. Indeed, that is even more evident when unknown properties are addressed, such as human behavior.

On the one hand, early representations of hybrid systems, {\it i.e.}, systems that mix continuous and discrete dynamics, consisted in instantiating discrete components and discretized continuous dynamics by using discrete models and considering non-modeled effects as uncertainties or perturbations exogenous to them, which can even be negligible. On the other hand, that approach is not suitable for complex ECPSs, which present intense nondeterminism, where uncertainties play a central role and might invalidate the entire system behavior.

ECPSs have been modeled through different structures that present different degrees of nondeterminism and balance between continuous and discrete dynamics, such as timed automata~\cite{Alur94}, hybrid automata~\cite{Alur93,Henzinger95}, and stochastic hybrid automata~\cite{Julius09,Ghosh05,Pola03}.

Timed automata employ finite automata with clocks, {\it i.e.}, real variables that follow continuous trajectories with constant slope. Hybrid automata generalize timed automata, also employing finite automata with real variables whose trajectories follow more general dynamical laws~\cite{Henzinger95}.  Timed automata and hybrid automata are deterministic model for hybrid systems that built the basis of HS representations, which were subdivided into different classes of hybrid automata, depending on the mapping functions ($r$ and $v$), {\it e.g.}, linear hybrid automata~\cite{Lafferriere99}, rectangular hybrid automata~\cite{Puri94,Henzinger95}, nonlinear hybrid automata~\cite{Henzinger95NL,Broucke1997}, integration graphs~\cite{Kesten1993}, and linear stochastic hybrid automata~\cite{Julius09}.

Rungger and Tabuada~\cite{Rungger16} and Tabuada {\it et al.}~\cite{Tabuada14} proposed a general and wide model for ECPS based on hybrid automata, providing the concept of robustness and important tools for robust stability analysis for ECPS.

@@

Given a class of ECPS's $\Omega$ and a set $\Phi$ of desired properties, a class of verification problems is called decidable, if there is an algorithm that is able to decide in a finite number of steps whether $\forall \mathcal{O}\in\Omega$ satisfies every property $\mathcal{L}\in \Phi$~\cite{Alur00}. Decidability is a key issue in ECPS verification, since there are uncountable states and complex dynamics to be considered, {\it i.e.}, hybrid automata and their variations (usually employed to model ECPS) are infinite-state machines~\cite{Henzinger95}. Generally, the properties that must hold for ECPS are related to safety, liveness, and reachability. In particular, model checking procedures have been employed to verify these properties of model programs even under undecidability conditions via approximations, {\it e.g.}, bounding the number of steps from the initial state, which leads to bounded model checking~\cite{Veanes09}.

>>>>

% As a general rule, do not put math, special symbols or citations
% in the abstract
\begin{abstract}
Dependency on the correct operation of embedded systems is rapidly growing, mainly due to their wide range of applications, such as micro-grids, automotive device control, health care, surveillance, mobile devices, telecommunications, internet of things, and consumer electronics. Their structures are becoming more and more complex and now require multi-core processors with scalable shared memory, signal-processing pipelines, and sophisticated software modules, in order to meet increasing computational power, flexibility demands, and adaptation to new scenarios and behaviors. Additionally, interaction with real-world entities and addition of modern communication capabilities further enhance the mentioned features and give rise to a new class of systems: embedded \& cyber-physical systems (ECPSs). As a consequence, reliability of ECPSs becomes a key issue during system development, which must be carefully addressed and assured. Normally, state-of-the-art verification methodologies for ECPSs generate test vectors (with constraints) and use assertion-based verification and high-level processor models, during simulation. Nonetheless, other additional challenges have been raised: the need for meeting time and energy constraints, handling concurrent software, dealing with platform restrictions, evaluating implementation-structure choices, validating operation logic, ensuring correct system behavior together with physical plants, providing compliance with target systems, incorporating knowledge about new problems and conditions, and supporting new software architectures and legacy designs. The present research discusses challenges, problems, and recent advances to ensure correctness and timeliness regarding ECPSs. Reliability issues, in the development of ECPSs, are then considered, as a prominent verification and synthesis application for achieving a \textit{correct-by-construction} design.
\end{abstract}

@@

The interaction between embedded software and physical processes created a different class of systems, which are complex, highly integrated, and present a mixture of continuous and discrete dynamics, named as \textit{hybrid systems} (HSs). Indeed, the embedded-system domain allied to the most recent communication revolution caused an intense integration between different and spread physical processes, which are called embedded \& cyber-physical systems (ECPSs). In particular, the presence of modern communication technologies ({\it e.g.}, internet of things - IoT) causes a revolution in terms of flexibility, scalability, and complexity of those systems and adds a novel class of problems to the challenges found in the embedded systems domain. For example, micro-grids, {\it i.e.}, small-scale electricity systems that gather different sources of distributed generation (DG) and loads, are emerging ECPSs, where reliability and carbon emission reduction are of paramount importance~\cite{xu15}. One may also notice that each DG element and the majority of current loads already employ embedded software and present several safety-critical requirements; however, the integration of both constitute an ECPS, where additional challenges related to synchronization, stability, control, communication, and reliability of the entire micro-grid arise.

@@

Figure~\ref{intelligent-product} shows some examples of ECPSs, which typically consist of a human-machine interface ({\it e.g.}, keyboard and display), a processing unit ({\it e.g.}, real-time computer system), and an instrumentation interface ({\it e.g.}, sensor, network, and actuator) that can be connected to some physical plant~\cite{Kopetz11}. Indeed, many current ECPSs, such as unmanned aerial vehicles (UAVs)~\cite{groza2015formal} and medical monitoring systems~\cite{Cordeiro09}, become interesting solutions only if they can reliably perform their target tasks. For instance, UAVs are a trend on military missions, due to the absence of pilots; however, an incorrect plan execution may cost civilian lives, which is unacceptable. In addition, wrong disease diagnosis or condition-evaluation reports have the potential to compromise patients' health and even many aspects of their lives, with serious consequences. On the one hand, portable ECPSs are capable of monitoring and identifying conditions that are very difficult for a physician, mainly when his contact with patients happens only in medical clinics. On the other hand, wrong or incomplete diagnostic data may delay necessary treatments or deteriorate patients' health. 
%
\begin{figure}[!t]
	\centering
	\includegraphics[scale=0.4]{figure1.jpg}
	\caption{An embedded system is part of a well-specified larger one (intelligent product).}
	\label{intelligent-product}
\end{figure}

Besides, when physical interaction with the real world is needed, which happens in ECPSs, additional care must be taken, mainly when human action is directly replaced, as in vehicle driving. Regarding the latter, even human-in-the-loop feedback control can be employed \cite{munir}, which then raises deeper concerns related to reliability of human behavior modeling and system implementation. Consequently, it is important to go beyond design correctness and also address behavior correctness, which may be performed by incorporating system models. Specifically, such models can be used to evaluate and even synthesize a given particular system, by ensuring that all needed functions are correctly implemented and the correct behavior is exhibited, {\it i.e.}, the system is indeed correct by its method of construction~\cite{Abate17}.

@@

The goals of the present article is to briefly present part of the established research and also clarify current trends, which are mainly driven by ECPSs, IoT, and ML. We present a brief discussion about models for ECPSs, in Section \ref{sec:model}, and then cover preliminaries on bounded model checking, \textit{k}-induction verification based on invariant inference, Craig interpolants, counterexample guided inductive synthesis, and system model incorporation, in Section~\ref{Preliminaries}. In Section~\ref{Verification-Challenges}, the main challenges regarding verification and synthesis of embedded systems are described, while Section~\ref{Research-Problem} tackles those challenges that are (partially) open in current published research. Section~\ref{achievements} then describes current achievements and future trends in verification and synthesis of ECPSs and Section \ref{Newapp} addresses new applications, which take into account the new trends ({\it e.g.}, ML) mentioned here. Finally, we conclude and describe future work, in Section~\ref{conclusions}.

@@

Automated formal verification has been employed to ensure correctness and reliability of ECPSs, during the last decade, as well as controller synthesis for ECPSs and HSs \textcolor{red}{add citations}.

An ECPS $S$ can be represented as a tuple $S=(X,X_{0},U,Y,r,v)$, where $X$ is a set of reachable states, $X_{0}$ is a set of initial states, $U$ is a set of inputs, $Y$ is the set of outputs, $r$ is a transition function such that $r:X\times U \rightarrow X$, and  $v$ is an output mapping function such that $v:X\times U \rightarrow Y$~\cite{Rungger16,Alur00,Girard11}.

The overall automated formal synthesis process for ECPSs and HSs involves three main steps: \textit{modeling}, which is related to obtaining a sound mathematical representation able to capture all relevant aspects of continuous and discrete dynamics and their interaction, as well as formulating properties and invariants that describe a desirable behavior, \textit{verification}, where a ECPS's behavior is checked through the proposed mathematical structure and compliance with a desired behavior is verified, and, finally, controller \textit{synthesis}, which produces a controller or finds invariant control laws that allow a ECPS to meet requirements and behave as desired, which is confirmed by a verification oracle.

@@

Models are largely employed in science and engineering areas, in order to analyze, predict, and modify the behavior of real-world systems. Specially, deterministic and stochastic models are the basis of many recent engineering advances. The former is uniquely determined by parameter values and previous states, while the latter presents inherent randomness, which is described by probability distributions~\cite{stochmodel}. Nonetheless, one can argue that such models are not enough to represent ECPSs, since interaction among different entities (software and hardware components) presents several events, which could not be deterministically or stochastically modeled; they should be better represented by nondeterministic approaches~\cite{leeCPS2}. Indeed, that is even more evident when unknown properties are addressed, such as human behavior.

On the one hand, early representations of hybrid systems, {\it i.e.}, systems that mix continuous and discrete dynamics, consisted in instantiating discrete components and discretized continuous dynamics by using discrete models and considering non-modeled effects as uncertainties or perturbations exogenous to them, which can even be negligible. On the other hand, that approach is not suitable for complex ECPSs, which present intense nondeterminism, where uncertainties play a central role and might invalidate the entire system behavior. Indeed, ECPSs have been modeled through different structures that present different degrees of nondeterminism and balance between continuous and discrete dynamics, such as timed automata~\cite{Alur94}, hybrid automata~\cite{Alur93,Henzinger95}, and stochastic hybrid automata~\cite{Julius09,Ghosh05,Pola03}.

Timed automata targets real-time systems and employs finite automata with clocks, {\it i.e.}, real variables that follow continuous trajectories (behaviors) with constant slope. Hybrid automata bridges digital processing and analog entities and consequently generalizes timed automata, by also employing finite automata with real variables, whose trajectories follow more general dynamical laws~\cite{Henzinger95}.  Timed automata and hybrid automata are deterministic models for hybrid systems that built the basis of HS representations, which were subdivided into different classes of hybrid automata, depending on mapping functions ($r$ and $v$), {\it e.g.}, linear hybrid automata~\cite{Lafferriere99}, rectangular hybrid automata~\cite{Puri94,Henzinger95}, nonlinear hybrid automata~\cite{Henzinger95NL,Broucke1997}, integration graphs~\cite{Kesten1993}, and linear stochastic hybrid automata~\cite{Julius09}.

Finally, Rungger and Tabuada~\cite{Rungger16} and Tabuada {\it et al.}~\cite{Tabuada14} proposed a general and wide model for ECPSs, which is based on hybrid automata, by providing the concept of robustness and important tools for robust stability analysis.

@@

Given a class of ECPSs $\Omega$ and a set $\Phi$ of desired properties, a class of verification problems is called decidable if there is an algorithm that is able to decide if a finite number of steps $\forall \mathcal{O}\in\Omega$ satisfy every property $\mathcal{L}\in \Phi$~\cite{Alur00}. Decidability is a key issue in ECPS verification, since there are uncountable states and complex dynamics to be considered, {\it i.e.}, hybrid automata and their variations (usually employed to model ECPSs) are infinite-state machines~\cite{Henzinger95}. Generally, the properties that must hold for ECPSs are related to safety, liveness, and reachability. In particular, model-checking procedures have been employed to verify those properties even under undecidability conditions via approximations, {\it e.g.}, bounding the number of steps from the initial state, which leads to bounded model checking~\cite{Veanes09}.
